### 2018-02-05 Segundo_Semestre/Seminario/Semana3/Presentacion.Rmd
inner circle and introduction

---
title: "Presentation"
author: "Mauricio Gonzalez"
date: "5/2/2018"
output: ioslides_presentation
bibliography: /Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib
---

# Introduction
## Intuition
- While playing a videogame or learning how to drive, human beings are forced to *make decisions* and to *actively intervene* in the world.
- This active *decision-making* process modifies the state of the world, which allows us to learn about *causal relations* that hold in our environment
- We ask ourselves how an *intelligent agent* can acquire and *use causal knowledge* while engaging with its environment to achieve some task.

# State of the Art
## Main groups and people
* Causality
    + Carnegie Mellon (Spirtes)
    + UC Los Angeles (Pearl)
    + MIT (Joshua Tenenbaum)
* Reinforcement Learning
    + Google DeepMind (David Silver, Alex Graves, V. Mnih, Shanahan, Garnelo)
    + Oxford (Nando De Freitas, Shimon Whiteson)
    + Cambridge (Ghahramani)
    + Alberta University (Richard Sutton, Csaba Szepesvári)
    + UC Berkeley (Pieter Abbeel)
    + Carnegie Mellon ()
    + Imperial College (Murray Shanahan, Marta Garnelo)
    + University College London

----

* Cognitive Science
    + Carnegie Mellon (David Danks)
    + MIT (Josh Tenenbaum)
    + UC Berkeley (Tom Griffiths)
    + Edinburgh University (Andy Clark)
    + Göttingen Institut für Psychologie (York Hagmayer)
    + Max Planck Institute for Human Development (Bjorn Meder)

## Outer Circle

Reference                  | Summary                                |
------------------------   |------------------------                |
[holland1986statistics]    | |
[@pearl1988probabilistic]  | Book on Networks for probabilistic reasoning|
[@spirtes2000causation]    | Classic General book on Causality      |
[@joyce1999foundations]    | Decision Theory and causal information |
[@pearl2009causality]      | Book on Causality and Graphical models |
[@koller2009probabilistic] | Book on Probabilistic Graphical Models |

----

Reference                    | Summary                                    |
-----------------------------|--------------------------------------------|
[@sutton1998reinforcement]   | Classic textbook on Reinforcement Learning |
[@van2012reinforcement]      | General review                             |
[@gershman2015reinforcement] | Reinforcement Learning in the Brain        |
[@li2017deep]                | Overview of Deep Reinforcement Learning    |

----

Reference            | Summary                                                                      |
---------------------|------------------------------------------------------------------------------|
[@krynski2007role]   |
[@danks2014unifying] | Cognitive functions can be modelled as operations on graphical models        |
[@lake2017building]  | Conditions for machines to be considered intelligent: Causal learning is one |

## Middle Circle
Reference | Summary |
----------|---------|
[@dawid2002influence] | |
[@hagmayer2009decision] | |
[@meder2009role] | |
[@meder2010observing] | |
[@kemp2010learning] | |
[@wellen2012learning] | |
[@mnih2013playing] | Introducing Deep Q-Network |
[@silver2014deterministic] | |
[@mnih2015human] | |
[@DBLP:journals/corr/MnihBMGLHSK16]
[@o2016pgq] | |

## Inner Circle

Reference | Summary |
----------|---------|
[@eberhardt2008causal] ||
[@eberhardt2012number] ||
[@hauser2012two] ||
[@hagmayer2013repeated] ||
[@ortega2014generalized] ||
[@alon2015online] ||
[@lattimoreNIPS2016] ||
[@garnelo2016towards] ||