### 2018-02-28 Segundo_Semestre/Seminario/Semana6/State_of_Art_presentation.Rmd
esqueleto presentacion

---
title: "State of the Art"
author: "Mauricio Gonzalez"
date: "28/2/2018"
output:
  beamer_presentation: default
  ioslides_presentation: default
  slidy_presentation: default
bibliography: /Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib
---

# Introduction
## Problem

# State of The Art

## Local updating

## Environment representations

## Sequential decision making

## Bayesian games