### 2018-03-05 Segundo_Semestre/Seminario/Semana6/State_of_Art_presentation.Rmd
presentacion final

---
title: "State of the Art"
author: "Mauricio Gonzalez"
date: "28/2/2018"
output:
  beamer_presentation:
   includes:
    in_header: header_pagenrs.tex
  ioslides_presentation: default
  slidy_presentation: default
bibliography: /Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib
---

# Introduction
## Problem
- Acquiring and using causal knowledge in a sequential decision-making process.

# State of The Art
## Environment representations
- In a series of papers, Seng-Beng Ho explains how to build a series of rules that encode causal relations found by exploration.
- In particular, in [@ho2016ground] they show how to discover (effective) causal relations in a stream of events where chronological order is important (e.g. lightning -> thunder)
- In [@ho2016reinforcement] they use the results of [@ho2014rapid] to build “scripts” that encode the local causal relations in order to choose actions.
- It is not an on-line procedure.

----

## Environment representations 2
- On the other side, it is shown by [@garnelo2016towards] that an agent can build on-line representations of the environment.
- They build a first-order logic representation which is then used to choose the next action.
- Not mainly focused on causal learning, but can be realized upon this learned representation.
- Small prototype designed ad-hoc.

----

## Sequential decision making
- A “bandit” is a slot machine in which you pull the lever and observe a reward.
- Used to mode sequential decision making: choose an action, observe reward.
- [@lattimoreNIPS2016] model how to choose inverventions in uncertain environment.
- They model each possible intervention as arms of a bandit.
- Fully specified causal model needed.

----

## Sequential decision making 2
- Thompson sampling: sample actions accordingly to (subjective) probability in action's optimality.
- [@ortega2014generalized] proposes Thompson sampling for causal induction.
- Agent samples parameter from beliefs, then chooses action as if parameter was true.
- Intervention and Bayesian updating.
- Not on-line

----

## Sequential decision making 2
- [@innes2018reasoning] shows how to incorporate unseen possibilities via environment exploration and expert advise. 
- Agent revises model of decision problem.
- Single-stage decision making.
- Extending results from [bramley2015staying] and [@albrecht2016exploiting] where set of variables was assumed to remain fixed.

----

## Bayesian Games and Belief updating
- [@bjorndahl2016bayesian] difference between intended and actual strategies
- [@mansour2015bayesian] & [mansour2016bayesian]: incentive-compatible exploration. When a “social planner” suggests actions.
- [@larrouy2017mindreading] psychological (but rational) construction of beliefs.
- [@amir2016endogenous] paying for extra information

# Idea of evaluation
## Max Expected Utility
- For a fully specified decision problem under uncertainty and a preference relations that satisfies Savage's axioms there exists a formal criterion for acting: max expected utility.
- If parameters are unknown, but agent has prior causal knowledge which will be updated in each round, the decisions used this learned model must be similar to MEU.
- Bayesian games with incomplete information allow for this belief updating.

----

## Simple toy example
- Patient come to a hospital and he can have eithe disease A or disease B.
- Medic can choose between doing nothing (N), give some pill (p) or sugery (s).
- If nothing is done, patient will die with probability 0.8.

----

- If medic choses surgery, patient can die, but if he survives and he had disease A, he will be cured with probability 0.5 or will not have any effect.
- If he had disease B, surgery saves him with probability 0.6

----

- If pill is given, he can die of allergic reaction, but if he survives and had disease A the pill cures him with probability 0.9.
- If he has disease B, pill has no effect.

----

- Now, suppose that this is a new disease, so the medic doesn't know exactly the probabilities.
- But he has previous causal info, from med school (e.g. pills cure anything always).
- He must update this info from experience in order to act as best as possible (bounded by MEU).
- Forget about ethics, this is only a mental experiment.

----