### 2018-12-05 WHY19/Draft/Paper.tex
Abstract y estructura

\def\year{2018}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai18}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
\pdfinfo{
/Title (A Guiding Principle for Causal Decision Problems)
/Author (M. Gonzalez-Soto, L.E. Sucar, H.J. Escalante)
}

\title{A Guiding Principle for Causal Decision Problems}
\author{M. Gonzalez-Soto, L.E. Sucar, H.J. Escalante\\
Address line
}

\begin{document}
\maketitle
\begin{abstract}
We define a Causal Decision Problem as a Decision Problem where available actions and outcomes are connected through the variables of a Causal Graphical Model \mathcal{G} and propose a solution criteria based on Pearl's Do-Calculus and the Expected Utility criteria for rational preferences. The implementation of this criteria leads to an on-line decision making procedure that has been shown to have similar performance to classic Reinforcement Learning algorithms while allowing for a causal model of an environment to be learned. 
\end{abstract}

\section{Introduction}
\noindent

\section{Rationality and Expected Utility}

\section{Optimal Policies}

\section{Causal Decision Problems}

\section{Proposed Solution}

\bibliography{Bibliografia.bib}
\bibliographystyle{aaai}
\end{document}