### 2020-09-10 SIGecom/Mauricio_Gonzalez_Profile.tex
profile for sigecom

\documentclass[acmsigecom]{acmtrans2m_nocategories}

\usepackage{enumitem,amsmath,amsfonts,setspace}
\usepackage{graphicx}
\usepackage{makeidx}
\makeindex

\usepackage{savesym}
\savesymbol{pdfbookmark}
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}
\restoresymbol{HR}{pdfbookmark}

\usepackage{etoolbox}
\makeatletter
\patchcmd{\l@section}
  {\hfil}
  {\leaders\hbox{\normalfont$\m@th\mkern \@dotsep mu\hbox{.}\mkern \@dotsep mu$}\hfill}
  {}{}
\makeatother

\makeatletter
\def\paragraph{\@startsection{paragraph}{4}%
  \z@\z@{-\fontdimen2\font}%
  {\normalfont\itshape}}
\makeatother

\markboth{}{Job Market Candidate Profiles 2021}

\title[Job Market Candidate Profiles 2021]{SIGecom Job Market Candidate Profiles 2021}

\begin{abstract} 
This is the sixth annual collection of profiles of the junior faculty job market candidates 
of the SIGecom community. The (number TBA) candidates for 2021 are listed alphabetically and 
indexed by research areas that define the interests of the community. The candidates can be 
contacted individually or via the moderated mailing list ecom-candidates2021@acm.org.\\

--Vasilis Gkatzelis and Jason Hartline
\vspace{-15pt}
\end{abstract}

\category{}{}{}
\setcounter{secnumdepth}{0}

\begin{document}

\maketitle

\tableofcontents
\newpage

% Replace CandidateFirstName and CandidateLastName with your first and last name.
% Choose some keywords (as many as you like, as long as it all fits in one line
% under the table of contents (which is, roughly, up to 70 characters).
% Replace the CandidateHomepage_URL and the CandidateCV_URL with the URL to your
% homepage and CV respectively. You can also add links to your Google scholar
% profile using the commented-out link. This is recommended, but not required.
\section[Mauricio Gonzalez-Soto\\ 
{\small keyword1, keyword2, keyword3, keyword4}]
{Mauricio Gonzalez-Soto 
(\href{https://www.researchgate.net/profile/Mauricio_Gonzalez-Soto}{Homepage},
\href{CandidateCV_URL}{CV}
%\href{CandidateGoogleScholar_URL}{Google Scholar}
)}

% Replace ThesisTitleCapitalized with the title of your thesis (if this is not finalized yet,
% then just choose some tentative one) and add the information of your advisor(s).
\textbf{Thesis:} ThesisTitleCapitalized

\vspace{3pt}
\noindent\textbf{Advisor:} Luis Enrique Sucar and Hugo J. Escalante, Instituto Nacional de Astrofísica Óptica y Electrónica.

% Distribute the remaining space however you like, but make sure that your whole profile fits
% within a page once you compile.
\vspace{3pt}
\noindent\textbf{Brief Biography:}
Mauricio received an Applied Mathematics Bachelors Degree from the Instituto Tecnológico Autónomo de México (ITAM) as well as a Masters Degree in Data Science from the same university, where he got awarded a Special Mention for each of his theses, he is currently studying a PhD in Computer Science under the supervision of Luis Enrique Sucar and Hugo J. Escalante.

% In your research summary, you can cite papers that are listed in the itemized environment below
% using static references: [1], [2],... Make sure that you cite the right paper if you later change
% the order in which your papers appear.
\vspace{3pt}
\noindent\textbf{Research Summary:}
Mauricio is mainly interested in Causality and its implications for Decision Making under Uncertainty

% In the itemized section below, you can list however many paper you like, but please use
% the same format that we have used in previous years for each publication. If you are not
% sure, just send an email to gkatz@drexel.edu
\vspace{3pt}
\noindent\textbf{Representative Papers:}
\begin{enumerate}[label={[\arabic*]}]
\item Causal Structure Learning using Random Graphs (Work in Progress: to be submitted to The 2020 NeurIPS Workshop on Causal Discovery and Causality-Inspired Machine Learning)\\
with Ivan Feliciano Avelino, L. Enrique Sucar and Hugo J. Escalante 
\item Von Neumann-Morgenstern and Savage Theorems for Causal Decision Making (Submitted to Artificial Intelligence Journal and oral presentation at New in ML Forum co-located with NeurIPS 2019) \href{https://arxiv.org/abs/1907.11752}{Link to Pre-print}\\
with L. Enrique Sucar and Hugo J. Escalante 
\item Causal Games and Causal Nash Equilibrium (1st Causal Reasoning Workshop at the Mexican International Conference on Artificial Intelligence 2019) \href{https://arxiv.org/abs/1910.06729}{Link to Pre-print}\\
with L. Enrique Sucar and Hugo J. Escalante 
\item Playing against Nature: causal discovery for decision making under uncertainty (Machine Learning for Causal Inference, Counterfactual Prediction and Autonomous Action (CausalML) Workshop at ICML 2018) \href{https://arxiv.org/abs/1807.01268}{Link to Pre-print}\\
with L. Enrique Sucar and Hugo J. Escalante

\end{enumerate}

\printindex

\end{document}